---
title: "07b_analysis"
author: "Sglatt"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages}
if (!require("survival")) {
  install.packages("survival")
  require("survival")
}
if (!require("survminer")) {
  install.packages("survminer")
  require("survminer")
}
if (!require("ggfortify")) {
  install.packages("ggfortify")
  require("ggfortify")
}
if (!require("dplyr")) {
  install.packages("dplyr")
  require("dplyr")
}
if (!require("psych")) {
  install.packages("psych")
  require("psych")
}
if (!require("pscl")) {
  install.packages("pscl")
  require("pscl")
}
if (!require("readxl")) {
  install.packages("readxl")
  require("readxl")
}
if (!require("pROC")) {
  install.packages("pROC")
  require("pROC")
}
if (!require("sjPlot")) {
  install.packages("sjPlot")
  require("sjPlot")
}
if (!require("extrafont")) {
  install.packages("extrafont")
  require("extrafont")
}
# font_import(prompt = FALSE)
# loadfonts(device = "win")
# fonts()
```

# Import the dataset created in 01_Data
```{r data and dir, eval=FALSE}
events_filtered <- read.csv("01_Output/Events_data_SG_2025-06-15.csv")

# Folder to save the output
dir.create("02_Output")
```

# Suicide descriptives for the sample
```{r describe}
# How many people had actual lifetime attempts?
events_filtered %>%
  count(cssrs_actattlt_bl, name = "Count", sort = FALSE) %>%
  mutate(Percent = round(100 * Count / sum(Count), 1))

# How many people had lifetime actual, aborted, and/or interrupted attempts?
events_filtered %>%
  count(attempt_lifetime, name = "Count", sort = FALSE) %>%
  mutate(Percent = round(100 * Count / sum(Count), 1))

# How many people had recent actual, aborted, and/or interrupted?
events_filtered %>%
  count(attempt_recent, name = "Count", sort = FALSE) %>%
  mutate(Percent = round(100 * Count / sum(Count), 1))

# How much baseline ideation did people have?
events_filtered %>%
  count(q06_ssi, name = "Count", sort = FALSE) %>%
  mutate(Percent = round(100 * Count / sum(Count), 1))

# How many people perceived access to method and opportunity for suicide?
events_filtered %>%
  count(q13_ssi, name = "Count", sort = FALSE) %>%
  mutate(Percent = round(100 * Count / sum(Count), 1))

# How many people prepared for suicide?
events_filtered %>%
  count(q16_ssi, name = "Count", sort = FALSE) %>%
  mutate(Percent = round(100 * Count / sum(Count), 1))

# How many people were in a high-risk hopeless group (9+)?
events_filtered %>%
  count(bhs_9, name = "Count", sort = FALSE) %>%
  mutate(Percent = round(100 * Count / sum(Count), 1))

# How many people had an attempt over the study period?
## Remove people who died with no emergent suicide attempt
## & people with < 340 days and no emergent suicide attempt
events_filtered_3 %>%
  count(att1_365, name = "Count", sort = FALSE) %>%
  mutate(Percent = round(100 * Count / sum(Count), 1))

table(events_filtered$att1_365)
table(events_filtered$natts_365 > 0)

# How many people more than 1 attempt over the study period?
## Remove people who died with no emergent suicide attempt and with an emergent suicide attempt(s)
events_filtered_2 %>%
  count(natts_365 > 1, name = "Count", sort = FALSE) %>%
  mutate(Percent = round(100 * Count / sum(Count), 1))

# What was the range of recurrence?
table(events_filtered$natts_365)

# How many people expected a future suicide attempt?
events_filtered %>%
  count(q15_ssi, name = "Count", sort = FALSE) %>%
  mutate(Percent = round(100 * Count / sum(Count), 1))

# How many people who expected a future suicide attempt had an emergent attempt over the study?
## Remove people who died with no emergent suicide attempt
## & people with < 340 days and no emergent suicide attempt

events_filtered_3 %>%
  count(q15_ssi, att1_365) %>%
  group_by(q15_ssi) %>%
  mutate(
    Percent = round(100 * n / sum(n), 1)
  ) %>%
  ungroup()

# Of the people who concretely expected to have/not have a suicide attempt, how many were *accurate*?

# Make an accuracy variable
events_filtered_3 <- events_filtered_3 %>%
  mutate(accuracy = case_when(
    q15_ssi == 0 & att1_365 == 0 ~ 1, # 1 if people expect not to make a suicide attempt and did not have one
    q15_ssi == 0 & att1_365 == 1 ~ 0, # 0 if people expect not to make a suicide attempt and did have one
    q15_ssi == 2 & att1_365 == 1 ~ 1, # 1 if people expect to make a suicide attempt and did have one
    q15_ssi == 2 & att1_365 == 0 ~ 0, # 0 if people expect to make a suicide attempt and did not have one
    TRUE ~ NA_real_
  ))

events_filtered_3[events_filtered_3$q15_ssi != 1, ] %>%
  count(q15_ssi_2, att1_365) %>%
  group_by(q15_ssi_2) %>%
  mutate(
    Percent = round(100 * n / sum(n), 1)
  ) %>%
  ungroup()

# Excluding everyone < 365 days
# Make an accuracy variable
events_filtered_4 <- events_filtered_4 %>%
  mutate(accuracy = case_when(
    q15_ssi == 0 & att1_365 == 0 ~ 1, # 1 if people expect not to make a suicide attempt and did not have one
    q15_ssi == 0 & att1_365 == 1 ~ 0, # 0 if people expect not to make a suicide attempt and did have one
    q15_ssi == 2 & att1_365 == 1 ~ 1, # 1 if people expect to make a suicide attempt and did have one
    q15_ssi == 2 & att1_365 == 0 ~ 0, # 0 if people expect to make a suicide attempt and did not have one
    TRUE ~ NA_real_
  ))

events_filtered_4[events_filtered_4$q15_ssi != 1, ] %>%
  count(q15_ssi_2, att1_365) %>%
  group_by(q15_ssi_2) %>%
  mutate(
    Percent = round(100 * n / sum(n), 1)
  ) %>%
  ungroup()
```

# Time to event descriptives?
```{r time to event}
# What was the mean time to first attempt?
## Remove people who died with no emergent suicide attempt
## & people with < 340 days and no emergent suicide attempt

summary(events_filtered_3[events_filtered_3$att1_365 == 1, ]$surv_365)
sd(events_filtered_3[events_filtered_3$att1_365 == 1, ]$surv_365)
table(events_filtered_3[events_filtered_3$att1_365 == 1, ]$surv_365)
nrow(events_filtered_3[events_filtered_3$att1_365 == 1, ])


# What was the mean time to first attempt when self-rated expectation == yes?
summary(events_filtered_3[events_filtered_3$att1_365 == 1 & events_filtered_3$q15_ssi_2 == 1, ]$surv_365)
sd(events_filtered_3[events_filtered_3$att1_365 == 1 & events_filtered_3$q15_ssi_2 == 1, ]$surv_365)
nrow(events_filtered_3[events_filtered_3$att1_365 == 1 & events_filtered_3$q15_ssi_2 == 1, ])

# What was the mean time to first attempt when self-rated expectation == no/unsure?
summary(events_filtered_3[events_filtered_3$att1_365 == 1 & events_filtered_3$q15_ssi_2 == 0, ]$surv_365)
sd(events_filtered_3[events_filtered_3$att1_365 == 1 & events_filtered_3$q15_ssi_2 == 0, ]$surv_365)
nrow(events_filtered_3[events_filtered_3$att1_365 == 1 & events_filtered_3$q15_ssi_2 == 0, ]$surv_365)
```

# Identify covariates with tests between emergent suicide attempt (absence vs. presence)
```{r cov}
## Remove people who died with no emergent suicide attempt
cov_1 <- fisher.test(events_filtered_1$att1_365, events_filtered_1$sex_demo)
cov_2 <- chisq.test(events_filtered_1$att1_365, events_filtered_1$race_demo)
cov_3 <- fisher.test(events_filtered_1$att1_365, events_filtered_1$latino_demo)
cov_4 <- chisq.test(events_filtered_1$att1_365, events_filtered_1$degree_demo)
cov_5a <- fisher.test(events_filtered_1$att1_365, events_filtered_1$prison_recoded)
cov_5b <- chisq.test(events_filtered_1$att1_365, events_filtered_1$prison_recoded)
cov_6 <- fisher.test(events_filtered_1$att1_365, events_filtered_1$prior_demo)
cov_7 <- chisq.test(events_filtered_1$att1_365, events_filtered_1$employ_demo)

report::report(cov_1)
report::report(cov_2)
report::report(cov_3)
report::report(cov_4)
report::report(cov_5b)
cov_5b$expected
cov_5b
report::report(cov_6)
report::report(cov_7)

cov_8 <- t.test(age_demo ~ att1_365, data = events_filtered_1)

report::report(cov_8)

# Look at a table of the significant covariate (justice involvement)
events_filtered_1 %>%
  count(att1_365, prison_recoded) %>%
  group_by(att1_365) %>%
  mutate(
    Percent = round(100 * n / sum(n), 1)
  ) %>%
  ungroup()

table(events_filtered_1$prison_recoded)
```

# Does a past-year attempt discern by emergent attempt?
```{r chi recent emergent}
## Remove people who died with no emergent suicide attempt
## & People with < 340 days and no emergent suicide attempt

chi_1 <- chisq.test(
  events_filtered_3$att1_365, # Suicide emergence (presence vs absence)
  events_filtered_3$attempt_recent # Recent suicide attempt (presence vs absence)
)
report::report(chi_1)

# Make a table
sjPlot::tab_xtab(
  var.row = events_filtered_3$att1_365,
  var.col = events_filtered_3$attempt_recent,
  title = "Contingency Table",
  show.row.prc = TRUE,
  show.col.prc = TRUE,
  show.summary = TRUE,
  statistics = "pearson",
  show.exp = TRUE,
  show.legend = TRUE,
  file = paste0("02_Output/Contingency_recent_emergent_", Sys.Date(), ".doc")
)
```

# Does perceived access to method and oppurtunity for suicide discern by emergent attempt?
```{r chi access emergent}
## Remove people who died with no emergent suicide attempt
## & People with < 340 days and no emergent suicide attempt

chi_2a <- chisq.test(
  events_filtered_3$att1_365, # Suicide emergence (presence vs absence)
  events_filtered_3$q13_ssi_2 # Perceived access to method/opportunity for suicide (no method and/or opportunity vs. yes)
)
report::report(chi_2a)
chi_2a
chi_2b <- chisq.test(
  events_filtered_3$att1_365, # Suicide emergence (presence vs absence)
  events_filtered_3$q13_ssi_3 # Perceived access to method/opportunity for suicide (no method vs. yes method and/or opportunity)
)
report::report(chi_2b)

# Make a table
sjPlot::tab_xtab(
  var.row = events_filtered_3$att1_365,
  var.col = events_filtered_3$q13_ssi_2,
  title = "Contingency Table",
  show.row.prc = TRUE,
  show.col.prc = TRUE,
  show.summary = TRUE,
  statistics = "pearson",
  show.exp = TRUE,
  show.legend = TRUE,
  file = paste0("02_Output/Contingency_access_emergent_", Sys.Date(), ".doc")
)
```

# Does self-rated expectation for suicide predict discern by emergent attempt?
```{r chi expect emergent}
## Remove people who died with no emergent suicide attempt
## & People with < 340 days and no emergent suicide attempt

chi_3 <- chisq.test(
  events_filtered_3$att1_365, # Suicide emergence (presence vs absence)
  events_filtered_3$q15_ssi_2 # Self-rated expectation (no/unsure vs. sure)
)
report::report(chi_3)

sjPlot::tab_xtab(
  var.row = events_filtered_3$att1_365,
  var.col = events_filtered_3$q15_ssi_2,
  title = "Contingency Table",
  show.row.prc = TRUE,
  show.col.prc = TRUE,
  show.summary = TRUE,
  statistics = "pearson",
  show.exp = TRUE,
  show.legend = TRUE,
  file = paste0("02_Output/Contingency_expect_emergent_", Sys.Date(), ".doc")
)
```

# Does baseline hopelessness discern by emergent attempt?
```{r chi expect emergent}
## Remove people who died with no emergent suicide attempt
## & People with < 340 days and no emergent suicide attempt

# Sum score
jmv::ttestIS(
  formula = bhs ~ # Hopelessness
    att1_365, # Suicide emergence (presence vs absence)
  data = events_filtered_3,
  vars = vars(bhs),
  welchs = TRUE,
  mann = TRUE,
  norm = TRUE,
  eqv = TRUE,
  meanDiff = TRUE,
  effectSize = TRUE,
  desc = TRUE
)

wilcox.test(bhs ~ att1_365, data = events_filtered_3)

# Dichotomous
chi_4 <- chisq.test(
  events_filtered_3$bhs_9,
  events_filtered_3$att1_365 # Self-rated expectation (no/unsure vs. sure)
)
report::report(chi_4)

# Make a table
sjPlot::tab_xtab(
  var.row = events_filtered_3$att1_365,
  var.col = events_filtered_3$bhs_9,
  title = "Contingency Table",
  show.row.prc = TRUE,
  show.col.prc = TRUE,
  show.summary = TRUE,
  statistics = "pearson",
  show.exp = TRUE,
  show.legend = TRUE,
  file = paste0("02_Output/Contingency_hope_emergent_", Sys.Date(), ".doc")
)
```

# Does preparation for suicide discern by emergent attempt?
```{r prep emergent}
## Remove people who died with no emergent suicide attempt
## & People with < 340 days and no emergent suicide attempt

chi_4 <- chisq.test(
  events_filtered_3$att1_365, # Suicide emergence (presence vs absence)
  events_filtered_3$q16_ssi_2 # preparation (none vs. some/finished)
)
report::report(chi_4)

sjPlot::tab_xtab(
  var.row = events_filtered_3$att1_365,
  var.col = events_filtered_3$q16_ssi_2,
  title = "Contingency Table",
  show.row.prc = TRUE,
  show.col.prc = TRUE,
  show.summary = TRUE,
  statistics = "pearson",
  show.exp = TRUE,
  show.legend = TRUE,
  file = paste0("02_Output/Contingency_prep_emergent_", Sys.Date(), ".doc")
)
```

```{r logit access}
logit_access_1 <- glm(
  q15_ssi_2 ~
    factor(bhs_9) +
    factor(attempt_recent) +
    factor(q13_ssi_2) +
    factor(q16_ssi_2),
  data = events_filtered,
  family = binomial()
)

summary(logit_access_1)
```

# Does expectation predict emergence (0 vs any)?
```{r logit emergent}
# Baseline predictors
logit_1a <- glm(
  att1_365 ~
    factor(arm_rec) + # Randomization condition
    factor(q15_ssi_2), # Self-rated expectation
  data = events_filtered_1,
  family = binomial()
)

# Adjust for recent suicidal behavior
logit_2a <- glm(
  att1_365 ~
    factor(arm_rec) +
    factor(q15_ssi_2) +
    (attempt_recent == 1), # Recent suicidal behavior
  data = events_filtered_1,
  family = binomial()
)

# Adjust for suicidal ideation
logit_3a <- glm(
  att1_365 ~
    factor(arm_rec) +
    factor(q15_ssi_2) +
    factor(q06_ssi),
  data = events_filtered_1,
  family = binomial()
)

# Adjust perceived access to method/opportunity for suicide
logit_4a <- glm(
  att1_365 ~
    factor(arm_rec) +
    factor(q15_ssi_2) +
    factor(q13_ssi_2),
  data = events_filtered_1,
  family = binomial()
)

# Adjust preparation for suicide
logit_5a <- glm(
  att1_365 ~
    factor(arm_rec) +
    factor(q15_ssi_2) +
    factor(q16_ssi_2),
  data = events_filtered_1,
  family = binomial()
)

# Adjust hopelessness - sum score
logit_6a1 <- glm(
  att1_365 ~
    factor(arm_rec) +
    factor(q15_ssi_2) +
    bhs,
  data = events_filtered_1,
  family = binomial()
)

# Adjust hopelessness - binary
logit_6a2 <- glm(
  att1_365 ~
    factor(arm_rec) +
    factor(q15_ssi_2) +
    factor(bhs_9),
  data = events_filtered_1,
  family = binomial()
)

logit_6a3 <- glm(
  att1_365 ~
    factor(arm_rec) +
    bhs,
  data = events_filtered_1,
  family = binomial()
)

# Initial model
summary(logit_1a)

# Adjusting for past-year suicidal behavior
summary(logit_2a)

# Adjusting for suicidal thoughts
summary(logit_3a)

# Adjusting for perceived access
summary(logit_4a)

# Adjusting for preparation for suicide
summary(logit_5a)

# Adjusting for hopelessness
summary(logit_6a1) # Sum score
summary(logit_6a2) # binary score
summary(logit_6a3) # Only hopelessness
```

# -- ROC curves
```{r roc}
models <- list(
  logit_1a,
  logit_2a,
  logit_3a,
  logit_4a,
  logit_5a,
  logit_6a1,
  logit_6a2
)

model_names <- c(
  "logit_1a",
  "logit_2a",
  "logit_3a",
  "logit_4a",
  "logit_5a",
  "logit_6a1",
  "logit_6a2"
)

roc_list <- list()
auc_values <- numeric(length(models))
auc_cis <- matrix(NA, nrow = length(models), ncol = 3)
colnames(auc_cis) <- c("l", "AUC", "u")
rownames(auc_cis) <- model_names

for (i in seq_along(models)) {
  pred_probs <- predict(models[[i]], type = "response")
  roc_obj <- roc(events_filtered_1$att1_365, pred_probs, ci = TRUE)
  roc_list[[model_names[i]]] <- roc_obj
  auc_values[i] <- auc(roc_obj)
  auc_ci <- ci.auc(roc_obj)
  auc_cis[i, ] <- auc_ci
  plot(roc_obj, main = paste("ROC Curve for", model_names[i]))
  cat(sprintf(
    "AUC for %s: %.3f (95%% CI: %.3f - %.3f)\n",
    model_names[i], auc_values[i], auc_ci[1], auc_ci[3]
  ))
}

auc_summary <- as.data.frame(auc_cis)
auc_summary
```

# Does expectation predict time to event?
```{r survival mods}
# Use full dataset

# Initial model
cox_1a <- coxph(
  Surv(surv_365, att1_365) ~
    factor(arm_rec) + # Randomization condition
    factor(q15_ssi_2), # Self rated expectation
  data = events_filtered
)

# Adjust for suicide behavior
cox_2a <- coxph(
  Surv(surv_365, att1_365) ~
    factor(arm_rec) + # Randomization condition
    factor(q15_ssi_2) + # Self rated expectation
    (attempt_recent == 1), # Recent suicidal behavior
  data = events_filtered
)

# Adjust for suicidal ideation
cox_3a <- coxph(
  Surv(surv_365, att1_365) ~
    factor(arm_rec) + # Randomization condition
    factor(q15_ssi_2) + # Self rated expectation
    factor(q06_ssi), # Suicidal ideation - brief, moderate, long
  data = events_filtered
)

# Adjust for perceived access to method and opportunity for suicide
cox_4a <- coxph(
  Surv(surv_365, att1_365) ~
    factor(arm_rec) + # Randomization condition
    factor(q15_ssi_2) + # Self rated expectation
    factor(q13_ssi_2), # Perceived access to method/opportunity for suicide
  data = events_filtered
)

# Adjust for preparation for suicide
cox_5a <- coxph(
  Surv(surv_365, att1_365) ~
    factor(arm_rec) + # Randomization condition
    factor(q15_ssi_2) + # Self rated expectation
    factor(q16_ssi_2), # Preparation
  data = events_filtered
)

# Adjust for hopelessness - sum score
cox_6a1 <- coxph(
  Surv(surv_365, att1_365) ~
    factor(arm_rec) + # Randomization condition
    factor(q15_ssi_2) + # Self rated expectation
    bhs, # hopelessness sum
  data = events_filtered
)

# Binary item
cox_6a2 <- coxph(
  Surv(surv_365, att1_365) ~
    factor(arm_rec) + # Randomization condition
    factor(q15_ssi_2) + # Self rated expectation
    factor(bhs_9), # hopelessness binary
  data = events_filtered
)

# Just hopelessness binary
cox_6a3 <- coxph(
  Surv(surv_365, att1_365) ~
    factor(arm_rec) + # Randomization condition
    factor(bhs_9), # hopelessness binary
  data = events_filtered
)

# Just hopelessness sum
cox_6a4 <- coxph(
  Surv(surv_365, att1_365) ~
    factor(arm_rec) + # Randomization condition
    bhs,
  data = events_filtered
)

summary(cox_1a)

# Adjusting for past-year suicidal behavior
summary(cox_2a)

# Adjusting for suicidal thoughts
summary(cox_3a)

# Adjusting for perceived access
summary(cox_4a)

# Adjusting for preparation for suicide
summary(cox_5a)

# Adjusting for hopelessness
summary(cox_6a1) # Sum score
summary(cox_6a2) # Binary score

# Test for improvements from the initial model -+- covariates
anova(cox_1a, cox_2a)
anova(cox_1a, cox_3a)
anova(cox_1a, cox_4a)
anova(cox_1a, cox_5a)
anova(cox_1a, cox_6a1)
anova(cox_1a, cox_6a2)
anova(cox_6a3, cox_6a2)


# Proportional hazards assumption
cox.zph(cox_1a)
cox.zph(cox_2a)
cox.zph(cox_3a)
cox.zph(cox_4a)
cox.zph(cox_5a)
cox.zph(cox_6a1)
cox.zph(cox_6a2)

# Summed hopelessness in model 5a violated the proportional hazards assumption.
# Re-fit the model with an hopelessness*time interaction

# Adjust for hopelessness with a time interaction
cox_6b1 <- coxph(
  Surv(surv_365, att1_365) ~
    factor(arm_rec) + # Randomization condition
    factor(q15_ssi_2) + # Self rated expectation
    bhs + # hopelessness
    tt(bhs), # hopelessness x time
  data = events_filtered,
  tt = function(x, t, ...) x * t # Define the time function for cox
)

summary(cox_6b1)
```

# Survival curve figure
```{r survival fig }
events_filtered <- events_filtered %>%
  mutate(q15_ssi_label = factor(q15_ssi_2, levels = c(0, 1), labels = c("No/Unsure", "Yes")))

cox_1a_fit <- survfit(Surv(surv_365, att1_365) ~ q15_ssi_label, data = events_filtered)

tiff(paste0("02_Output/Survival_plot_", format(Sys.Date(), "%Y-%m-%d"), ".tiff"), units = "in", width = 8, height = 5, res = 300)

autoplot(
  cox_1a_fit,
  conf.int = FALSE,
  censor.shape = "*",
  censor.size = 5,
  surv.size = 1.1,
  facets = TRUE,
  ncol = 2
) +
  labs(
    x = "Time to suicide attempt",
    y = "Survival",
    color = "Expectation of a future suicide attempt"
  ) +
  coord_cartesian(ylim = c(0.25, 1)) +
  theme_minimal(base_size = 14, base_family = "Times New Roman") +
  theme(
    panel.grid.major = element_line(color = "grey90"),
    panel.grid.minor = element_line(color = "grey95"),
    panel.background = element_rect(fill = "white"),
    strip.background = element_rect(fill = "white"),
    legend.position = "bottom"
  )
dev.off()
```

# Do people with recent behavior anticipate having suicidal behavior at higher rates?
```{r chi recent expect}
# With the full sample
chi_5a <- chisq.test(
  events_filtered$q15_ssi_2, # Self-rated expectation (no/unsure vs. sure)
  events_filtered$attempt_recent
) # recent suicide attempt (presence vs. absence)
report::report(chi_5a)

# Make sure chi is appropriate
chi_5a$expected

# only with the people who made concrete expectations (exclude "unsure" expectations)
chi_5b <- chisq.test(
  x = events_filtered[events_filtered$q15_ssi != 1, ]$q15_ssi_2, # Self-rated expectation (unsure vs. sure)
  y = events_filtered[events_filtered$q15_ssi != 1, ]$attempt_recent # recent suicide attempt (presence vs. absence)
)
report::report(chi_5b)
chi_5b$expected

# How many in the sample with concrete expectations had an emergent suicidal behavior?
table(events_filtered[events_filtered$q15_ssi != 1, ]$attempt_recent)

# Make tables
# -- full sample
sjPlot::tab_xtab(
  var.row = events_filtered$q15_ssi_2,
  var.col = events_filtered$attempt_recent,
  title = "Contingency Table",
  show.row.prc = TRUE,
  show.col.prc = TRUE,
  show.summary = TRUE,
  statistics = "pearson",
  show.exp = TRUE,
  show.legend = TRUE,
  file = paste0("02_Output/Contingency_full_recent_expect_", Sys.Date(), ".doc")
)

# -- subsample
sjPlot::tab_xtab(
  var.row = events_filtered[events_filtered$q15_ssi != 1, ]$q15_ssi_2,
  var.col = events_filtered[events_filtered$q15_ssi != 1, ]$attempt_recent,
  title = "Contingency Table",
  show.row.prc = TRUE,
  show.col.prc = TRUE,
  show.summary = TRUE,
  statistics = "pearson",
  show.exp = TRUE,
  show.legend = TRUE,
  file = paste0("02_Output/Contingency_sub_recent_expect_", Sys.Date(), ".doc")
)
```

# Do people with method/opportunty anticipate having suicidal behavior at higher rates?
```{r chi access expect}
# with the full sample
chi_6a <- chisq.test(
  events_filtered$q15_ssi_2, # Self-rated expectation (no/unsure vs. sure)
  events_filtered$q13_ssi_2
) # Perceived access to method/opportunity for suicide (no method and/or opportunity vs. yes)
report::report(chi_6a)
chi_6a$expected

# only with the people who made concrete expectations (exclude "unsure" expectations)
chi_6b <- chisq.test(
  x = events_filtered[events_filtered$q15_ssi != 1, ]$q15_ssi_2, # Self-rated expectation (unsure vs. sure)
  y = events_filtered_1[events_filtered_1$q15_ssi != 1, ]$q13_ssi_2 # Perceived access to method/opportunity for suicide (no method and/or opportunity vs. yes)
)
chi_6b
chi_6b$expected # Make sure chi is appropriate

# How many in the sample with concrete expectations percieved access to method and opportunity for suicide
table(events_filtered_1[events_filtered_1$q15_ssi != 1, ]$q13_ssi_2)

# Make tables
# -- full sample
sjPlot::tab_xtab(
  var.row = events_filtered$q15_ssi_2,
  var.col = events_filtered$q13_ssi_2,
  title = "Contingency Table",
  show.row.prc = TRUE,
  show.col.prc = TRUE,
  show.summary = TRUE,
  statistics = "pearson",
  show.exp = TRUE,
  show.legend = TRUE,
  file = paste0("02_Output/Contingency_full_access_expect_", Sys.Date(), ".doc")
)

# -- subsample
sjPlot::tab_xtab(
  var.row = events_filtered[events_filtered$q15_ssi != 1, ]$q15_ssi_2,
  var.col = events_filtered[events_filtered$q15_ssi != 1, ]$q13_ssi_2,
  title = "Contingency Table",
  show.row.prc = TRUE,
  show.col.prc = TRUE,
  show.summary = TRUE,
  statistics = "pearson",
  show.exp = TRUE,
  show.legend = TRUE,
  file = paste0("02_Output/Contingency_sub_access_expect_", Sys.Date(), ".doc")
)
```

# Do people with prep for suicide anticipate having suicidal behavior at higher rates?
```{r chi prep expect}
# with the full sample
chi_7a <- chisq.test(
  events_filtered$q15_ssi_2, # Self-rated expectation (no/unsure vs. sure)
  events_filtered$q16_ssi_2 # preparation for suicide (no vs. some/completed-finished)
)
report::report(chi_7a)
chi_7a$expected

# Make tables
# -- full sample
sjPlot::tab_xtab(
  var.row = events_filtered$q15_ssi_2,
  var.col = events_filtered$q16_ssi_2,
  title = "Contingency Table",
  show.row.prc = TRUE,
  show.col.prc = TRUE,
  show.summary = TRUE,
  statistics = "pearson",
  show.exp = TRUE,
  show.legend = TRUE,
  file = paste0("02_Output/Contingency_full_prep_expect_", Sys.Date(), ".doc")
)
```

# Are people who anticipate suicidal behavior (vs unsure/no) more hopeless?
```{r t t chi hope expect}
# Full sample, continuous score
jmv::ttestIS(
  formula = bhs ~ q15_ssi_2,
  data = events_filtered,
  vars = vars(bhs),
  welchs = TRUE,
  mann = TRUE,
  norm = TRUE,
  eqv = TRUE,
  meanDiff = TRUE,
  effectSize = TRUE,
  desc = TRUE
)

wilcox.test(bhs ~ q15_ssi_2, data = events_filtered)

# Full sample, binary item
chi_8a <- chisq.test(
  events_filtered$q15_ssi_2, # Self-rated expectation (no/unsure vs. sure)
  events_filtered$bhs_9
) # Perceived access to method/opportunity for suicide (no method and/or opportunity vs. yes)
report::report(chi_8a)
chi_8a$expected

# only with the people who made concrete expectations (exclude "unsure" expectations)
# Sum score
jmv::ttestIS(
  formula = bhs ~ q15_ssi_2,
  data = events_filtered[events_filtered$q15_ssi != 1, ],
  vars = vars(bhs),
  welchs = TRUE,
  mann = TRUE,
  norm = TRUE,
  eqv = TRUE,
  meanDiff = TRUE,
  effectSize = TRUE,
  desc = TRUE
)

wilcox.test(bhs ~ q15_ssi_2, data = events_filtered, subset = !is.na(accuracy))

# binary item
chi_8b <- chisq.test(
  events_filtered[events_filtered$q15_ssi != 1, ]$q15_ssi_2, # Self-rated expectation (no/unsure vs. sure)
  events_filtered[events_filtered$q15_ssi != 1, ]$bhs_9
)
report::report(chi_8b)
chi_8b$expected

# Make tables
# -- full sample
sjPlot::tab_xtab(
  var.row = events_filtered$q15_ssi_2,
  var.col = events_filtered$bhs_9,
  title = "Contingency Table",
  show.row.prc = TRUE,
  show.col.prc = TRUE,
  show.summary = TRUE,
  statistics = "pearson",
  show.exp = TRUE,
  show.legend = TRUE,
  file = paste0("02_Output/Contingency_full_hope_expect_", Sys.Date(), ".doc")
)

# -- subsample
sjPlot::tab_xtab(
  var.row = events_filtered[events_filtered$q15_ssi != 1, ]$q15_ssi_2,
  var.col = events_filtered[events_filtered$q15_ssi != 1, ]$bhs_9,
  title = "Contingency Table",
  show.row.prc = TRUE,
  show.col.prc = TRUE,
  show.summary = TRUE,
  statistics = "pearson",
  show.exp = TRUE,
  show.legend = TRUE,
  file = paste0("02_Output/Contingency_sub_hope_expect_", Sys.Date(), ".doc")
)
```

# Do people with recent behavior have more *accurate* predictions?
```{r chi recent accuracy}
# How many people did and did not expect a suicide concretely?
table(
  events_filtered_1$q15_ssi, # Self-rated expectations
  events_filtered_1$attempt_recent
) # Recent suicidal behaviors

# How many of those concrete expectations were accurate?
table(
  events_filtered$q15_ssi, # Self-rated expectations
  events_filtered$accuracy
) # Accuracy of self-rated expectations

# How many people in the sample with concrete expectations had recent suicidal behavior?
table(events_filtered[!is.na(events_filtered$accuracy), ]$attempt_recent)

chi_9 <- chisq.test(
  events_filtered$accuracy, # Accuracy of self-rated expectations (yes vs. no)
  events_filtered$attempt_recent
) # Recent suicidal behavior (presence vs. absence)
report::report(chi_9)

chi_9$expected # Make sure chi is appropriate

sjPlot::tab_xtab(
  var.row = events_filtered[events_filtered$q15_ssi != 1, ]$accuracy,
  var.col = events_filtered[events_filtered$q15_ssi != 1, ]$attempt_recent,
  title = "Contingency Table",
  show.row.prc = TRUE,
  show.col.prc = TRUE,
  show.summary = TRUE,
  statistics = "pearson",
  show.exp = TRUE,
  show.legend = TRUE,
  file = paste0("02_Output/Contingency_recent_accuracy_", Sys.Date(), ".doc")
)
```

# Do people with access to method/oppurtunity have more *accurate* predictions?
```{r chi access accuracy}
chi_10 <- chisq.test(
  events_filtered$accuracy, # Accuracy of self-rated expectations (yes vs. no)
  events_filtered$q13_ssi_2
) # Perceived access to method/opportunity for suicide (no method and/or opportunity vs. yes)
report::report(chi_10)
chi_10$expected # Make sure chi is appropriate

# Table of accuract and perceived access
events_filtered[events_filtered$q15_ssi != 1, ] %>%
  count(accuracy, q13_ssi_2) %>%
  group_by(accuracy) %>%
  mutate(
    Percent = round(100 * n / sum(n), 1)
  ) %>%
  ungroup()

sjPlot::tab_xtab(
  var.row = events_filtered[events_filtered$q15_ssi != 1, ]$accuracy,
  var.col = events_filtered[events_filtered$q15_ssi != 1, ]$q13_ssi_2,
  title = "Contingency Table",
  show.row.prc = TRUE,
  show.col.prc = TRUE,
  show.summary = TRUE,
  statistics = "pearson",
  show.exp = TRUE,
  show.legend = TRUE,
  file = paste0("02_Output/Contingency_access_accuracy_", Sys.Date(), ".doc")
)
```

# Are people who are more hopeless have more *accurate* predictions?
```{r t chi hopelessness accuracy}
# Sum score
jmv::ttestIS(
  formula = bhs ~ accuracy,
  data = events_filtered,
  vars = vars(bhs),
  welchs = TRUE,
  mann = TRUE,
  norm = TRUE,
  eqv = TRUE,
  meanDiff = TRUE,
  effectSize = TRUE,
  desc = TRUE
)

wilcox.test(bhs ~ accuracy, data = events_filtered, subset = !is.na(accuracy))

# Binary item
table(
  events_filtered$accuracy, # Accuracy of self-rated expectations (yes vs. no)
  events_filtered$bhs_9 # Binary hopelessness
)

chi_11 <- chisq.test(
  events_filtered$accuracy, # Accuracy of self-rated expectations (yes vs. no)
  events_filtered$bhs_9 # Binary hopelessness
)
report::report(chi_11)
chi_11$expected # Make sure chi is appropriate

sjPlot::tab_xtab(
  var.row = events_filtered[events_filtered$q15_ssi != 1, ]$accuracy,
  var.col = events_filtered[events_filtered$q15_ssi != 1, ]$bhs_9,
  title = "Contingency Table",
  show.row.prc = TRUE,
  show.col.prc = TRUE,
  show.summary = TRUE,
  statistics = "pearson",
  show.exp = TRUE,
  show.legend = TRUE,
  file = paste0("02_Output/Contingency_hope_accuracy_", Sys.Date(), ".doc")
)
```

# Do people with preparation for suicide have more *accurate* predictions?
```{r chi prep accuracy}
chi_12 <- chisq.test(
  events_filtered$accuracy, # Accuracy of self-rated expectations (yes vs. no)
  events_filtered$q16_ssi_2 # Binary hopelessness
)
report::report(chi_12)
chi_12$expected # Make sure chi is appropriate

# Table of accuracy and perceived access
events_filtered[events_filtered$q15_ssi != 1, ] %>%
  count(accuracy, q16_ssi_2) %>%
  group_by(accuracy) %>%
  mutate(
    Percent = round(100 * n / sum(n), 1)
  ) %>%
  ungroup()

sjPlot::tab_xtab(
  var.row = events_filtered[events_filtered$q15_ssi != 1, ]$accuracy,
  var.col = events_filtered[events_filtered$q15_ssi != 1, ]$q16_ssi_2,
  title = "Contingency Table",
  show.row.prc = TRUE,
  show.col.prc = TRUE,
  show.summary = TRUE,
  statistics = "pearson",
  show.exp = TRUE,
  show.legend = TRUE,
  file = paste0("02_Output/Contingency_prep_accuracy_", Sys.Date(), ".doc")
)
```

# Does recent suicidal behavior relate to actual suicidal behavior in the subsample with accurate predictions?
```{r chi recent emergent}
chi_13 <- chisq.test(
  events_filtered[events_filtered$accuracy == 1, ]$att1_365,
  events_filtered[events_filtered$accuracy == 1, ]$attempt_recent
)
report::report(chi_13)
chi_13$expected

fisher_13 <- fisher.test(
  events_filtered[events_filtered$accuracy == 1, ]$att1_365,
  events_filtered[events_filtered$accuracy == 1, ]$attempt_recent
)
fisher_13
```

# Among who do not expect to attempt suicide, does recent suicidal behavior distinguish between those who do versus don’t engage in suicidal behavior?
```{r}
chi_14 <- chisq.test(
  events_filtered[events_filtered$q15_ssi_2 == 0, ]$att1_365,
  events_filtered[events_filtered$q15_ssi_2 == 0, ]$q13_ssi_2
)
chi_14
chi_14$observed
```


# Woodpile 

# Does self-rated expectation predict presence of an emergent suicide attempt (and recurrence among people with one)
```{r hurdle mods}
## Remove people who died with no emergent suicide attempt and with an emergent suicide attempt(s)

# baseline predictors
hurdle_1a <- hurdle(
  natts_365 ~
    factor(arm_rec) + # Randomization condition
    factor(q15_ssi_2), # Self-rated expectation
  data = events_filtered_1,
  dist = "negbin"
)

# Adjust for recent suicidal behavior
hurdle_2a <- hurdle(
  natts_365 ~
    factor(arm_rec) +
    factor(q15_ssi_2) +
    factor(attempt_recent), # Recent suicidal behavior
  data = events_filtered_1,
  dist = "negbin"
)

# Adjust for suicidal ideation
hurdle_3a <- hurdle(
  natts_365 ~
    factor(arm_rec) +
    factor(q15_ssi_2) +
    factor(q06_ssi),
  data = events_filtered_1,
  dist = "negbin"
)

# Adjust perceived access to method/opportunity for suicide
hurdle_4a <- hurdle(
  natts_365 ~
    factor(arm_rec) +
    factor(q15_ssi_2) +
    factor(q13_ssi_2),
  data = events_filtered_1,
  dist = "negbin"
)

# Adjust hopelessness
hurdle_5a <- hurdle(
  natts_365 ~
    factor(arm_rec) +
    factor(q15_ssi_2) +
    bhs,
  data = events_filtered_1,
  dist = "negbin"
)

# Initial model
summary(hurdle_1a)

# Adjusting for past-year suicidal behavior
summary(hurdle_2a)

# Adjusting for suicidal thoughts
summary(hurdle_3a)

# Adjusting for perceived access
summary(hurdle_4a)

# Adjusting for hopelessness
summary(hurdle_5a)
```

# Does expectation predict overall frequency?
```{r zin}
## Remove people who died with no emergent suicide attempt and with an emergent suicide attempt(s)

# Check for overdisperson (variance > mean)
mean(events_filtered_2$natts_365, na.rm = TRUE)
var(events_filtered_2$natts_365, na.rm = TRUE)

# Fit a Poisson GLM for formal dispersion test
poisson_model <- glm(
  natts_365 ~ factor(q15_ssi_2),
  family = poisson,
  data = events_filtered_2
)

dispersiontest(poisson_model)

# Test if a zero-inflated negative binomial is better then negative binomial
nb_model <- MASS::glm.nb(
  natts_365 ~
    factor(q15_ssi_2),
  data = events_filtered_2
)

zinb_model <- zeroinfl(
  natts_365 ~
    factor(q15_ssi_2) | 1,
  dist = "negbin",
  data = events_filtered_2
)

AIC(nb_model, zinb_model)
BIC(nb_model, zinb_model)

# Initial model
nb_model_1a <- MASS::glm.nb(
  natts_365 ~
    factor(q15_ssi_2) + # Self-rated expectation
    factor(arm_rec), # Treatment condition
  data = events_filtered_2
)

# Adjust for past-year suicidal behavior
nb_model_2a <- MASS::glm.nb(
  natts_365 ~
    factor(q15_ssi_2) + # Self-rated expectation
    factor(arm_rec) + # Treatment condition
    factor(attempt_recent), # Recent suicidal behavior
  data = events_filtered_2
)

# Adjust for suicidal ideation
nb_model_3a <- MASS::glm.nb(
  natts_365 ~
    factor(q15_ssi_2) + # Self-rated expectation
    factor(arm_rec) + # Treatment condition
    factor(q06_ssi), # Suicidal ideation - brief, moderate, long
  data = events_filtered_2
)

# Adjust for perceived access to method and opportunity for suicide
nb_model_4a <- MASS::glm.nb(
  natts_365 ~
    factor(q15_ssi_2) + # Self-rated expectation
    factor(arm_rec) + # Treatment condition
    factor(q13_ssi_2), # Perceived access to suicide methods/opportunity
  data = events_filtered_2
)

# Adjust for hopelessness
nb_model_5a <- MASS::glm.nb(
  natts_365 ~
    factor(q15_ssi_2) + # Self-rated expectation
    factor(arm_rec) + # Treatment condition
    bhs,
  data = events_filtered_2
)

# Initial model
summary(nb_model_1a)

# Adjusting for past-year suicidal behavior
summary(nb_model_2a)

# Adjusting for suicidal thoughts
summary(nb_model_3a)

# Adjusting for perceived access
summary(nb_model_4a)

# Adjusting for hopelessness
summary(nb_model_5a)
```
